---
title: "Assignment 1, Part 2"
author: "Blanka Zana & Riley Anthony"
thing: "12 December 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 1, Part 2

## Exercise 1) Preliminary Data Exploration

Describe the participant samples in the dataset (e.g. by diagnosis, age, etc.). Do you think the two groups are well balanced? If not, what do you think was the reason?

```{r}
library(dplyr)
library(lme4)
library(ggplot2)
library(lmtest)
library(pacman)
library(nlme)
library(lmerTest)
p_load(MuMIn,stats,pcaMethods)
setwd("~/dánia/Study/Semeser III/ExpMeth III/Assignment 1/Assignment-1")
thing = read.csv2("DONE.csv")
thing[1] = NULL

thing$SUBJ = as.factor(thing$SUBJ)
summary(thing$Ethnicity[unique(thing$SUBJ)])
summary(thing$Ethnicity[thing$Diagnosis =="TD"])
summary(thing$Ethnicity[thing$Diagnosis =="ASD"])
summary(thing$Gender[thing$Diagnosis == "ASD"][thing$VISIT==3], na.rm = T)
summary(thing$Gender[thing$Diagnosis=="ASD"])

v1 = subset(thing, VISIT == 1)
summary(v1$Gender[v1$Diagnosis =="ASD"])
summary(v1$Gender[v1$Diagnosis =="TD"])

summary(v1$Ethnicity[v1$Diagnosis =="ASD"])
summary(v1$Ethnicity[v1$Diagnosis =="TD"])


#1)
mean(thing$Age[thing$Diagnosis == "TD"],na.rm=T)
mean(thing$Age[thing$Diagnosis == "ASD"],na.rm=T)

ggplot(thing,aes(x=VISIT, y=CHI_MLU, color=Diagnosis)) + geom_boxplot() + geom_smooth(method="lm",se=TRUE) + facet_wrap(~VISIT)

#2)
ggplot(thing,aes(Diagnosis,Age))+geom_violin()

#3)
thing$SUBJ=as.factor(thing$SUBJ)
#depicting the results throughout all visits of all children together
debasement = ggplot(thing,aes(VISIT,CHI_MLU,color=SUBJ))+geom_line()+facet_wrap(~Diagnosis)+theme(legend.position="none") 
debasement
#same, but separating TD and ASD kids results
debasement1 = ggplot(thing,aes(VISIT,CHI_MLU,color=SUBJ))+geom_line()+theme(legend.position="none") 
debasement1

#4)
ggplot(thing,aes(VISIT,tokens_CHI))+geom_point()+geom_smooth(method="lm",se=TRUE)+facet_wrap(~Diagnosis)

```

### Exercise 2) Children learning language: the effects of time and ASD
Describe linguistic development in TD and ASD children in terms of Mean Length of Utterance (MLU)?

```{r}
#plot first

ggplot(thing,aes(VISIT,CHI_MLU)) + geom_point() + geom_smooth(method="lm",se=T)

flabbergasted = lmer(CHI_MLU ~ VISIT + (1|SUBJ) + (1|Ethnicity) + (1|Gender), thing)
summary(flabbergasted)
anova(flabbergasted)

PLOT = ggplot(thing, aes(x= Age, y = CHI_MLU)) +
  geom_point(aes())+
  facet_wrap(~Diagnosis)+
  geom_smooth(method = "lm")

PLOT

#ii)

#plot first
ggplot(thing,aes(Diagnosis, CHI_MLU)) + geom_boxplot()
nebulous = lmer(CHI_MLU ~ Diagnosis + (1|SUBJ) + (1|Ethnicity) + (1|Gender), thing)
summary(nebulous)
anova(nebulous)

#iii)
#plot first

ggplot(thing,aes(VISIT,CHI_MLU)) + geom_point() + geom_smooth(method="lm", se= T) + facet_wrap(~Diagnosis)

serendipity = lmer(CHI_MLU~ Diagnosis+VISIT+Age + (1|SUBJ) + (1|Gender), thing)
summary(serendipity)
anova(serendipity)
r.squaredGLMM(serendipity)
plot(serendipity)

base = lm(CHI_MLU ~ Diagnosis, thing)

##############this one
timediagmodel = lmer(CHI_MLU ~ VISIT * Diagnosis + (1+VISIT|SUBJ), thing)
summary(timediagmodel)
r.squaredGLMM(timediagmodel)

###############this one

timemodel = lmer(CHI_MLU ~ VISIT + (1+VISIT|SUBJ), thing)
r.squaredGLMM(timemodel)
summary(timemodel)

################this one
nullmodel = lmer(CHI_MLU ~ 1 + (1+VISIT|SUBJ), thing)
r.squaredGLMM(nullmodel)
summary(nullmodel)

anova(timediagmodel,timemodel,nullmodel)

nullmodel = lmer(CHI_MLU ~ 1 + (1+VISIT|SUBJ), thing)
r.squaredGLMM(nullmodel)
anova(nullmodel,timemodel)
anova(nullmodel,timemodel,timediagmodel)
summary(nullmodel)

ggplot(thing, aes(VISIT,CHI_MLU)) + geom_violin() + facet_wrap(~Diagnosis)

```

### Exercise 3) Child directed speech as a moving target
Describe how parental use of language changes over time in terms of MLU. What do you think is going on?

```{r}

parentime = lmer(MOT_MLU~VISIT+(1+VISIT|SUBJ), thing, REML = F)
summary(parentime)
r.squaredGLMM(parentime)

parentimediag = lmer(MOT_MLU~VISIT+Diagnosis+(1+VISIT|SUBJ), thing, REML = F)
summary(parentimediag)
r.squaredGLMM(parentimediag)

nullparent = lmer(MOT_MLU~1+(1+VISIT|SUBJ), thing, REML = F)

anova( parentimediag,nullparent)

ggplot(thing, aes(Age,MOT_MLU))+geom_point()+geom_smooth(method = "lm", se=T,fullrange=T)+facet_wrap(~Diagnosis)+theme(legend.position="none")
```

### Exercise 4) Looking into "individual differences" (demographic, clinical or cognitive profiles)
Describe your strategy to select the best models (how did you choose the variables to include?) and send the code to Riccardo and Celine.

```{r}
best=lmer(CHI_MLU~VISIT*Diagnosis+verbalIQ+tokens_CHI++types_CHI+(1+VISIT|SUBJ), thing)
summary(best)
r.squaredGLMM(best)
anova(best)

other = lmer(CHI_MLU~VISIT+ADOS+(1+VISIT|SUBJ), thing)
summary(other)
r.squaredGLMM(other)

other2 = lmer(CHI_MLU~MOT_MLU*Diagnosis+ADOS+(1+VISIT|SUBJ), thing)
summary(other2)
r.squaredGLMM(other2)

beste=lmer(CHI_MLU~VISIT*Diagnosis+verbalIQ+tokens_CHI+types_CHI+(1+VISIT|SUBJ), thing)
summary(beste)
r.squaredGLMM(beste)

```
