---
title: "Assignment 5, Part 1"
author: "Blanka Zana & Riley Anthony"
date: "12 December 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


setwd("C:/Users/tamec/Desktop/Programming/Git/ExpMeth Assignment 8 Data")
library(pacman)
p_load(lmerTest, metafor)

```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive patterns of pitch mean and pitch sd in schizophrenia? Report how many papers report quantitative estimates, your method to analyze them, the estimated effect size of the difference (mean effect size and standard error for pitch mean, same for pitch sd) and forest plots representing it. 

```{r}
#Load data

df = read.csv(file = "A5Data.csv", sep = ";")

#Calculate cohens D
  #mean
df = escalc('SMD', n1i = SampleSizeSchizo, n2i = SampleSizeContros, m1i = PitchMeanSchizo, m2i = PitchMeanControls, sd1i = PitchMeanSchizoSD, sd2i = PitchMeanControlsSD, data = df)
  #SD
df1 = escalc('SMD', n1i = SampleSizeSchizo, n2i = SampleSizeContros, m1i = PitchSDSchizo, m2i = PitchSDControls, sd1i = PitchSDSchizoSD, sd2i = PitchSDControlsSD, data = df)


m1.1 = rma(yi, vi, data = df, slab = Article)
forest(m1.1)
summary(m1.1)


m1.2 = rma(yi, vi, data = df1, slab = Article)
forest(m1.2)
summary(m1.2)


##Question 1 
#14 were included in the meta analysis, all reporting quantitative estimates. Using metafor (citation) Cohen's d was calculated for studies including means and standard deviation. Resulting in a total of 4 studies. This resulted in the following forrest plot:

#Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.
#making a df with my own data
Idat = data.frame(Article = "my_data", Year = "2017", SampleSizeSchizo = 57, SampleSizeContros = 59, PitchMeanControls = 135.0317, PitchMeanControlsSD = 25.61366, PitchMeanSchizo = 149.079, PitchMeanSchizoSD = 24.02952, PitchSDControls = 47.7987, PitchSDControlsSD = 17.90346, PitchSDSchizo = 57.83132, PitchSDSchizoSD = 36.78883)
  #mean
Idat = escalc('SMD', n1i = SampleSizeSchizo, n2i = SampleSizeContros, m1i = PitchMeanSchizo, m2i = PitchMeanControls, sd1i = PitchMeanSchizoSD, sd2i = PitchMeanControlsSD, data = Idat)
  #SD
Idat1 = escalc('SMD', n1i = SampleSizeSchizo, n2i = SampleSizeContros, m1i = PitchSDSchizo, m2i = PitchSDControls, sd1i = PitchSDSchizoSD, sd2i = PitchSDControlsSD, data = Idat) # adds yi and vi (yi= cohens D, vi = variance of cohens D)


  #adding my own data to the forrest plot
    #mean
df2 = rbind(df, Idat)
m2 = rma(yi, vi, data = df2, slab = Article)
forest(m2)
    #SD
df3 = rbind(df1, Idat1)
m2.1 = rma(yi, vi, data = df3, slab = Article)
forest(m2.1)


```

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

```{r}

#Adding our own analysis from assignment 3 result in the following forrest plot:
#Comparing the two forrest plots, we observe that our own study change the estimated effect size. (how to argument here?)

#Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.
  #mean
funnel(m1.1, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
m1.1
funnel(m2, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
m2
regtest(m2) #no systematic relation (linear assumption)
ranktest(m2) # no systematic relation (ordinal assumption)
inf = influence(m2) #martinez change the result significantly 
plot(inf)


  #SD
funnel(m1.2, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
funnel(m2.1, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
regtest(m2.1) #no systematic relation (linear assumption)
ranktest(m2.1) # no systematic relation (ordinal assumption)
inf = influence(m2.1) #martinez change the result significantly 
plot(inf)
#?# how is this calculated? 

#without cohen 2014
df3.1 = subset(df3, Article != "Cohen et al. 2014")
m2.2 = rma(yi, vi, data = df3.1, slab = Article)
forest(m2.2)
funnel(m2.2, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
regtest(m2.2) #no systematic relation (linear assumption)
ranktest(m2.2) # no systematic relation (ordinal assumption)


```

3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

```{r}

#tau^2 (estimated amount of total heterogeneity): 0.0846 (SE = 0.1053)
#I^2 (total heterogeneity / total variability):   58.88%


```
## Tips on the process to follow:

- Download the data on all published articles analyzing pitch in schizophrenia: https://www.dropbox.com/s/pmgw8wzc308so1p/Data.csv?dl=0
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2